---
title: "Owrasse_Underwater_Behavior"
format: pdf
editor: visual
---

# Quick Intro

This quarto document contains the code used to visualize data from UW_Nest_Behavioral_Cleaned. The goals of this project is to understand what factors influence the mating decisions that female Ocellated wrasse make. What makes *Symphodus ocellatus* (the ocellated wrasse) interesting is their 3 different male reproductive tactics. Additionally, there are large amounts of mate choice copying from the females, so much so that a male is up significantly more likely to spawn with a female if she observes another female with him ([Alonzo 2008](https://www.sciencedirect.com/science/article/pii/S0003347208000225)).

The data used was collected live underwater by SCUBA divers in Corsica, Fr, in a 10 minute period.

# Tidy Code

First run our library list (Install first if you don't have these packages).

```{r}
#Install and library important packages
#install.packages('ggplot2')
#install.packages('here')
#install.packages('tidyverse')
#install.packages('readxl')
library(tidyverse)
library(here)
library(ggplot2)
library(readxl)
library(dplyr)
```

Importing the data sheet

```{r}
behavior_sheet <- read_excel("UW_Nest_Behavioral_Sheets_Cleaned.xlsx")
head(behavior_sheet)
```

Separating the date column and placing it near the original Date column for filtering ease later on.

```{r eval=FALSE}
behavior_sheet_date <- as.Date(behavior_sheet$Date)
behavior_sheet$Date <- behavior_sheet_date
behavior_sheet_sep <- behavior_sheet %>%
  mutate(Year = year(Date)) %>%
  mutate(Month = month(Date)) %>%
  mutate(Day = day(Date)) %>%
  relocate(Year, .after = Date) %>%
  relocate(Month, .after = Year) %>%
  relocate(Day, .after = Month)
head(behavior_sheet_sep)
```
