---
title: "Data_Analysis_Code"
subtitle: Analysis Code and Narrative Descriptions
format: pdf
editor: visual
---

# Goals

## Broad question

What social factors are important in female mate choice?

## Specific question(s)

Do interactions with certain male reproductive tactics increase a female's propensity to spawn at a nest?

Does nesting male courtship behavior increase the propensity for a female to visit or spawn with that male?

How does nesting male aggression influence female visits and spawning?

### This document

The purpose of this document is to catch you up o the progress of our research on mate choice! It will help you visualize what we've been doing and what we've found! Please ask questions, there are no wrong questions we are all learning together!

# Tidying code/Wrangling

Here we are downloading and getting all the important libraries we will need for the code. Then we are going to download the raw data. Then we fix some of the columns so we can use them for analysis, especially the date column!

```{r}
#Install and library important packages
#install.packages('ggplot2')
#install.packages('ggeffects')
#install.packages('RColorBrewer')
#install.packages('here')
#install.packages('MASS')
#install.packages('tidyverse')
#install.packages('janitor')
#install.packages('readxl')
#install.packages('brms')
#install.packages('ProbBayes')
library(ProbBayes)
library(ggeffects)
library(tidyverse)
library(janitor)
library(here)
library(ggplot2)
library(readxl)
library(dplyr)
library(brms)
library(RColorBrewer)
library(tibble)
library(MASS)

behavior_sheet <- read_excel("UW_Nest_Behavioral_Sheets_Cleaned.xlsx")

behavior_sheet_date <- as.Date(behavior_sheet$Date)
behavior_sheet$Date <- behavior_sheet_date
behavior_sheet_sep <- behavior_sheet %>%
  mutate(Year = year(Date)) %>%
  mutate(Month = month(Date)) %>%
  mutate(Day = day(Date)) %>%
  relocate(Year, .after = Date) %>%
  relocate(Month, .after = Year) %>%
  relocate(Day, .after = Month)
head(behavior_sheet_sep)

#Fix SAT submission to NM from chr to num
sat_sub <- as.numeric(behavior_sheet_sep$`SAT submission to NM`) #Creates a vector with the same entries but as numbers
behavior_sheet_sep$`SAT submission to NM` <- sat_sub #put it back into the tibble
#Do the same for SNK Submission to SAT
snk_sub <- as.numeric(behavior_sheet_sep$`SNK submission to SAT`)
behavior_sheet_sep$`SNK submission to SAT` <- snk_sub
#Same for NM aggression to SAT
NM_aggsat <- as.numeric(behavior_sheet_sep$`NM aggression to SAT`)
behavior_sheet_sep$`NM aggression to SAT` <- NM_aggsat

behavior_sheet_cln <- behavior_sheet_sep |>
  #Uses the Janitor package to standardize the names
  clean_names()

#Turn the year variable into a character for ease of comparison later on
year_date <- as.character(behavior_sheet_cln$year) #Creates a vector with years as chr
behavior_sheet_cln$year <- year_date #creates a new variable for ease of graphing later on
```

# Visualization

## Female Spawns as a function of Female Visits

This is just to help us visualize what sort of processes are occurring at the nest.

```{r}
fem.ts.fem.v <- behavior_sheet_cln |>
  ggplot(aes(total_female_visits, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot
  geom_smooth(method = 'glm', se = F, method.args = list(family = 'poisson'), aes(color = "Poisson")) + #Plots a Poisson model onto the data
  scale_color_manual(name = 'Model', breaks = "Poisson", values = c('Poisson' = 'darkgreen'), guide = 'legend') +
  theme_classic() + #Sets the plot visuals
  labs(x = 'Female Visits (count)', y = 'Females that spawned (count)', title = "How many females spawned as a function of total Female Visits") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.ts.fem.v
```

## Female Visits/Spawns as a function of Average Number of Sneakers

### Total Female visits as a function of Average Number of sneakers

```{r}
fem.v.avg.snk <- behavior_sheet_cln |>
  ggplot(aes(average_snk, total_female_visits)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot 
  geom_smooth(se = F, color = 'darkgreen') + #Adds a trendline for better visuals
  theme_classic() + #Sets the plot visuals
  labs(x = 'Average Number of Sneakers (count)', y = 'Females that visited (count)', title = "How many females visited the nest as a function of Average Number of Sneakers") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.v.avg.snk
  
```

### Total Females spawned as a function of Average Number of sneakers

```{r}
fem.s.avg.snk <- behavior_sheet_cln |>
  ggplot(aes(average_snk, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot 
  geom_smooth(se = F, color = 'darkgreen') + #Adds a trendline for better visuals
  theme_classic() + #Sets the plot visuals
  labs(x = 'Average Number of Sneakers (count)', y = 'Females that spawned (count)', title = "How many females spawned at the nest as a function of Average Number of Sneakers") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.s.avg.snk
```

## Female that spawns as a function of being brought to the nest by the Nesting Male

```{r}
fem.s.bnm <- behavior_sheet_cln |>
  mutate('nm.brought.spawns'= (female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)) |>
  ggplot(aes(nm.brought.spawns, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot 
  geom_smooth(method = 'glm.nb', se = F, color = 'darkgreen') + #Plots a Negative Binomial model onto the data
  theme_classic() + #Sets the plot visuals
  labs(x = 'Spawns when brought by the Nesting Male (count)', y = 'Females that spawned (count)', title = "How many females spawned as a function of being brought by the Nesting Male") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.s.bnm
```

## Female Spawns as a function of being brought to the nest by the Satellite Male

```{r}
fem.s.bsat <- behavior_sheet_cln |>
  mutate('sat.brought.spawns'= (female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)) |>
  ggplot(aes(sat.brought.spawns, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot 
  geom_smooth(method = 'glm.nb', se = F, color = 'darkgreen') + #Plots a Negative Binomial model onto the data
  theme_classic() + #Sets the plot visuals
  labs(x = 'Spawns when brought by the Satellite Male (count)', y = 'Females that spawned (count)', title = "How many females spawned as a function of being brought by the Satellite Male") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.s.bsat
```

## Female Spawns as a function of being brought to the nest by the Satellite Male and passed to the Nesting Male

```{r}
fem.s.bsat.pnm <- behavior_sheet_cln |>
  mutate('bsat.pnm.spawns'= (female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)) |>
  ggplot(aes(bsat.pnm.spawns, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot 
 geom_smooth(method = 'glm.nb', se = F, color = 'darkgreen') + #Plots a Negative Binomial model onto the data
  theme_classic() + #Sets the plot visuals
  labs(x = 'Spawns when brought by the Satellite Male but passed (count)', y = 'Females that spawned (count)', title = "How many females spawned as a function of being brought by the Satellite Male but passed off") + #Sets the axis and title labels
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fem.s.bsat.pnm
```

## Female Visits/Spawns as a function of Nesting Male Courtship

### Total Female Visits as a function of Nesting Male Courtship

```{r}
fem.v.nmcourt <- behavior_sheet_cln |>
  ggplot(aes(nm_courtship, total_female_visits, color = average_snk)) + #gives plot dimensions
  geom_point() + #scatterplot
  geom_smooth(method ='glm.nb', se = F, color = 'darkgreen') + #Plots a Negative Binomial model onto the data
  theme_classic()+
  labs(x = 'How many times nesting male displayed courtship behavior (count)', y = 'Females that visited (count)', title = 'How many females visited as a function of nesting male courtship events', color = 'Avg # sneakers') #Sets axis and title

plot(fem.v.nmcourt)

```

### Total Female Spawns as a function of Nesting Male Courtship

```{r}
fem.s.nmcourt <- behavior_sheet_cln |>
  ggplot(aes(nm_courtship, total_female_spawning, color = average_snk)) + #gives plot dimensions
  geom_point() + #scatterplot
  geom_smooth(method ='glm.nb', se = F, color = 'darkgreen') + #Plots a Negative Binomial model onto the data
  theme_classic()+
  labs(x = 'How many times nesting male displayed courtship behavior (count)', y = 'Females that visited (count)', title = 'How many females visited as a function of nesting male courtship events', color = 'Avg # sneakers') #Sets axis and title

plot(fem.s.nmcourt)
```

## Female visits/spawning as a function of nesting male aggression

### Female visits as a function of nesting male aggression

```{r}
fvisits.nmagg <- behavior_sheet_cln |>
  #creates a variable that is a total of all NM agression
  mutate('total_nm_agg' = nm_aggression_to_sat + nm_aggression_to_snk + nm_aggression_to_other_nm + nest_defence_nm_to_other_spp) |>
  #Plotting previous variable to female visits
  ggplot(aes(total_nm_agg, total_female_visits)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot
  geom_smooth() + #adds a trendline for the plot
  labs(x = 'Total nesting male aggression behaviors',
       y = 'Total female visits',
       title = 'How many females visited as a function of nesting male aggression') + #Makes labels
  theme_classic() + #Sets the plot visuals
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

fvisits.nmagg
```

### Female spawning as a function of nesting male aggression

```{r}
f.spawns.nmagg <- behavior_sheet_cln |>
  #creates a variable that is a total of all NM agression
  mutate('total_nm_agg' = nm_aggression_to_sat + nm_aggression_to_snk + nm_aggression_to_other_nm + nest_defence_nm_to_other_spp) |>
  #Plotting previous variable to female visits
  ggplot(aes(total_nm_agg, total_female_spawning)) + #gives ggplot dimensions
  geom_point() + #makes the plot a scatterplot
  geom_smooth() + #adds a trendline for the plot
  labs(x = 'Total nesting male aggression behaviors',
       y = 'Total female spawns',
       title = 'How many female spawns as a function of nesting male aggression') + #Makes labels
  theme_classic() + #Sets the plot visuals
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle

f.spawns.nmagg
```

# Models

### Generalized Linear Models

### Female Visits/Spawns as a function of Nesting Male Courtship

#### Female visits as a function of nesting male courtship

```{r}
fv.nm.court <-
  brm(data = behavior_sheet_cln, #Give the model the tibble
      family = poisson(link = 'log'), #Tell model what type of distribution
      total_female_visits ~ nm_courtship, #Specifying the model
      iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fv.nm.court') #saves it as an output

print(fv.nm.court, digits=4)
plot(fv.nm.court)
```

#### Female spawning as a function of nesting male courtship

```{r}
fs.nm.court <-
  brm(data = behavior_sheet_cln, #Give the model the tibble
      family = poisson(link = 'log'), #Tell model what type of distribution
      total_female_spawning ~ nm_courtship, #Specifying the model
      iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fs.nm.court') #saves it as an output

print(fs.nm.court, digits=4)
plot(fs.nm.court)
```

### Female visits/spawning as a function of nesting male aggression

#### Female visits as a function of nesting male aggression

```{r}
fvisits.nmagg <- behavior_sheet_cln |>
  #creates a variable that is a total of all NM agression
  mutate('total_nm_agg' = nm_aggression_to_sat + nm_aggression_to_snk + nm_aggression_to_other_nm + nest_defence_nm_to_other_spp)

fvisits.nmagg.mdl1 <-
  brm(data = fvisits.nmagg, #Give the model the tibble
      family = negbinomial(link = 'log'), #Tell model what type of distribution
      total_female_visits ~ total_nm_agg, #Specifying the model
      iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fvisits.nmagg.mdl1') #saves it as an output

print(fvisits.nmagg.mdl1, digits = 4)
plot(fvisits.nmagg.mdl1)
```

#### Female spawning as a function of nesting male aggression

```{r}
f.spawns.nmagg <- behavior_sheet_cln |>
  #creates a variable that is a total of all NM agression
  mutate('total_nm_agg' = nm_aggression_to_sat + nm_aggression_to_snk + nm_aggression_to_other_nm + nest_defence_nm_to_other_spp)

f.spawns.nmagg.mdl1 <-
  brm(data = f.spawns.nmagg, #Give the model the tibble
      family = negbinomial(link='log'), #Tell model what type of distribution
      total_spawns ~ total_nm_agg, #Specifying the model
      iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/f.visits.nmagg.mdl1') #saves it as an output

print(f.spawns.nmagg.mdl1, digits = 4) #Gives a summary of the model
plot(f.spawns.nmagg.mdl1) #Plots the chains
```

### Female spawns as a function of Her Interaction Pathway to the nest

#### Note:

While our data is based on counts, usually making model distributions poisson or negative binomial, we will be analyzing our data counts as binomial. This is because we are trying to see if the "interaction path", what social interaction the female had as she visited the nest, changes the probability of whether or not the female will spawn at that nest. Thus we can consider the number of female visits as the number of trials, with the probability of spawning for each interaction path being the probability.

#### Normalizing the pathways

By normalizing our pathways to the total number of visits, we are making each variable as a fraction of visits for each pathways, i.e. fraction of females that spawned which were brought by the SAT, fraction of females that spawned which were brought by the SAT and passed to the NM. This allows us to understand the probability of a certain pathway leading to spawning success.

```{r}
sheet.normalized <- behavior_sheet_cln |>
  #Normalize the different pathways by the total number of females that spawned
  mutate(
    #Nesting male brings pathway
    'fspawn.nm.b' = ((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/total_female_visits),
    #Satellite male brings pathway
    'fspawn.sat.b' = ((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/total_female_visits),
    #Satellite male passes to Nesting Male pathway
    'fspawn.sat.b.p.nm' = ((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/total_female_visits)
  )
  
```

#### Model 1 Just Nesting Male

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.1 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.nm.b),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.1') #saves it as an output

#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.1)
#Gives us the summary of the model
print(fem.spawns.path.mdl.1, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.1)
```

#### Model 2 Just Satellite

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.2 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.sat.b),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.2') #saves it as an output

#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.2)
#Gives us the summary of the model
print(fem.spawns.path.mdl.2, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.2)
```

#### Model 3 Just Satellite pass to Nesting Male

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.3 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.sat.b.p.nm),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.3') #saves it as an output

#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.3)
#Gives us the summary of the model
print(fem.spawns.path.mdl.3, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.3)
```

#### Model 4, Just Nesting Male and Just Satellite

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.4 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.nm.b + fspawn.sat.b),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.4') #saves it as an output

#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.4)
#Gives us the summary of the model
print(fem.spawns.path.mdl.4, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.4)
```

#### Model 5, Just Nesting Male and Just Satellite pass to Nesting Male

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.5 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.nm.b + fspawn.sat.b.p.nm),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.5') #saves it as an output

#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.5)
#Gives us the summary of the model
print(fem.spawns.path.mdl.5, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.5)
```

#### Model 6, Just Satellite and Just Satellite pass to Nesting Male

```{r}
sheet.normalized.rmv <- sheet.normalized |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


fem.spawns.path.mdl.6 <- sheet.normalized.rmv |>
  brm(data = sheet.normalized.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + fspawn.sat.b + fspawn.sat.b.p.nm),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 2000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/fem.spawns.path.mdl.6') #saves it as an output


#Gives a plot of the model's predictions
pp_check(fem.spawns.path.mdl.6)
#Gives us the summary of the model
print(fem.spawns.path.mdl.6, digits = 4)
#Plots the Markov chains
plot(fem.spawns.path.mdl.6)
```

#### Compare the models

```{r}
waic(fem.spawns.path.mdl.1)
waic(fem.spawns.path.mdl.2)
waic(fem.spawns.path.mdl.3)
waic(fem.spawns.path.mdl.4)
waic(fem.spawns.path.mdl.5)
waic(fem.spawns.path.mdl.6)
```

Based on the waic values, model 5 which is our model with the Nesting male bringing females and Satellite passing to Nesting male pathways is the best model given our data. This makes biological sense, the better a Nesting male and Satellite pair are the more likely a nest is going to be efficient, chasing sneakers and bringing females!

## More Aligned Models

While the previous model is interesting, it isn't quite telling us what we want to know. Our previous model was inferring that the total number of females spawning were best predicted by the females that spawned when the nesting male brought a female or when a Satellite passed the female to the nesting male. What we really want to know is if there is a difference in female spawning decision based on which 'pathway' she was introduced to the nest. Thus we need to make the model a little more complex.

Goal: Given how a female visits, does it change the likelihood of spawning (in each pathway). Does that then correlate to overall nest success? Are there other influencing factors (number of sneakers, +/- nesting male and satellite male interactions?)

Let's visualize the pathways.

```{r}
pathways.visits <- behavior_sheet_cln |>
    #Remove nests where there are data inconsistencies
  filter_out(female_visits_by_nm > total_female_visits) |>
  filter_out(female_visit_by_sat > total_female_visits) |>
  filter_out(female_visits_by_sat_passed_to_nm > total_female_visits)|>
  filter_out(female_visit_by_snk > total_female_visits)|>
  #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = sum(female_visits_by_nm)/sum(total_female_visits),
    'Satellite Male' = sum(female_visit_by_sat)/sum(total_female_visits),
    'Satellite Male Passes' = sum(female_visits_by_sat_passed_to_nm)/sum(total_female_visits),
    'Sneaker Male' = sum(female_visit_by_snk)/sum(total_female_visits)
  ) |>
  #Make the tibble only include these variables
  dplyr::select(`Nesting Male`:`Sneaker Male`) |>
  #Remove all the duplicates
  distinct() |>
  #Adjust the table to make the variables go into one coloumn
  pivot_longer(`Nesting Male`:`Sneaker Male`,
               names_to = 'Pathway',
               values_to = 'Proportion') |>
  #Graph it
  ggplot(aes(Pathway, Proportion, fill = Pathway)) +
  #Makes the graph a bar graph
  geom_col() +
  labs(x = 'Pathway', 
       y = 'Proportion of total Visits',
       title = "The Propotion of each Visitation Pathway for Female Visits",) + #edits axis and title labels
  theme_classic() + #Sets the plot visuals
  #Adds a color scheme to the plot
  scale_fill_manual(values = c('Nesting Male' = '#065143', 'Satellite Male' = '#db5a42', 'Satellite Male Passes' = '#7A9B76','Sneaker Male' = '#1481BA')) +
  theme(plot.title = element_text(hjust = 0.5)) + #Moves title to the middle
  theme(legend.position = "none") # Removes all legends
  
pathways.visits
```

Now let's plot the success rate of each pathway:

```{r}
pathways.success.rate <- behavior_sheet_cln |>
  #Remove nests where there are data inconsistencies
  filter_out((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/(female_visits_by_nm) > 1) |>
  filter_out((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/(female_visit_by_sat) > 1) |>
  filter_out((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/(female_visits_by_sat_passed_to_nm) > 1) |>
  filter_out((female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)/(female_visit_by_snk) > 1)|>
  #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = sum(female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/sum(female_visits_by_nm),
    'Satellite Male' = sum(female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/sum(female_visit_by_sat),
    'Satellite Pass' = sum(female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/sum(female_visits_by_sat_passed_to_nm),
    'Sneaker Male' = sum(female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)/sum(female_visit_by_snk)
  ) |>
  #Make the tibble only include these variables
  dplyr::select(`Nesting Male`:`Sneaker Male`) |>
  #Remove all the duplicates
  distinct() |>
  #Adjust the table to make the variables go into one coloumn
  pivot_longer(`Nesting Male`:`Sneaker Male`,
               names_to = 'Pathway',
               values_to = 'Success Rate') |>
  #Graph it
  ggplot(aes(Pathway, `Success Rate`, fill = Pathway)) +
  #Makes the graph a bar graph
  geom_col() +
  labs(x = 'Pathway', 
       y = 'Success rate',
       title = "The Success Rate of each Visitation Pathway",) + #edits axis and title labels
  theme_classic() + #Sets the plot visuals
  #Adds a color scheme to the plot
  scale_fill_manual(values = c('Nesting Male' = '#065143', 'Satellite Male' = '#db5a42', 'Satellite Pass' = '#7A9B76','Sneaker Male' = '#1481BA')) +
  theme(plot.title = element_text(hjust = 0.5)) + #Moves title to the middle
  theme(legend.position = "none") # Removes all legends
  
pathways.success.rate
```

The presence of the sneaker male rate could be because of secondary satellites being counted as sneakers

Now let's visualize how often these pathway successes occur in total at a nest.

```{r}
pathways.success.rate.total <- behavior_sheet_cln |>
  #Remove nests where there are data inconsistencies
  filter_out((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/(total_female_spawning)> 1 ) |>
  filter_out((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/(total_female_spawning) > 1) |>
  filter_out((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/(female_visits_by_sat_passed_to_nm) > 1) |>
  filter_out((female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)/(total_female_spawning) > 1)|>
  #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = sum(female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/sum(total_female_spawning),
    'Satellite Male' = sum(female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/sum(total_female_spawning),
    'Satellite Male Pass' = sum(female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/sum(total_female_spawning),
    'Sneaker Male' = sum(female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)/sum(total_female_spawning)
  ) |>
  #Make the tibble only include these variables
  dplyr::select(`Nesting Male`:`Sneaker Male`) |>
  #Remove all the duplicates
  distinct() |>
  #Adjust the table to make the variables go into one coloumn
  pivot_longer(`Nesting Male`:`Sneaker Male`,
               names_to = 'Pathway',
               values_to = 'Success Rate') |>
  #Graph it
  ggplot(aes(Pathway, `Success Rate`, fill = Pathway)) +
  #Makes the graph a bar graph
  geom_col() +
  labs(x = 'Pathway', 
       y = 'Success rate',
       title = "The Proportion of Total Females that spawned for each Visitation Pathway",) + #edits axis and title labels
  theme_classic() + #Sets the plot visuals
  #Adds a color scheme to the plot
  scale_fill_manual(values = c('Nesting Male' = '#065143', 'Satellite Male' = '#db5a42', 'Satellite Male Pass' = '#7A9B76','Sneaker Male' = '#1481BA')) +
  theme(plot.title = element_text(hjust = 0.5)) + #Moves title to the middle
  theme(legend.position = "none") # Removes all legends
  
pathways.success.rate.total
```

Let's visualize if certain pathways lead to females spawning more

```{r}
pathways.success.rate.total <- behavior_sheet_cln |>
  #Remove nests where there are data inconsistencies
  filter_out((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm) > (total_spawns_nm_brought + total_spawns_nm_brought_sat_only_spawn)) |>
  filter_out((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn) > (total_spawns_sat_brought + total_spawns_sat_brought_sat_only_spawn)) |>
  filter_out((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn) > (total_spawns_sat_passed_to_nm + total_spawns_sat_passed_to_nm_but_sat_only_spawn))|>
  filter_out((female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn) > (total_spawns_snk_brought + total_spawns_snk_brought_snk_only_spawn))|>
  #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = sum(total_spawns_nm_brought + total_spawns_nm_brought_sat_only_spawn)/sum(female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm),
    'Satellite Male' = sum(total_spawns_sat_brought + total_spawns_sat_brought_sat_only_spawn)/sum(female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn),
    'Satellite Pass' = sum(total_spawns_sat_passed_to_nm + total_spawns_sat_passed_to_nm_but_sat_only_spawn)/sum(female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn),
    'Sneaker Male' = sum(total_spawns_snk_brought + total_spawns_snk_brought_snk_only_spawn)/sum(female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)
  ) |>
  #Make the tibble only include these variables
  dplyr::select(`Nesting Male`:`Sneaker Male`) |>
  #Remove all the duplicates
  distinct() |>
  #Adjust the table to make the variables go into one coloumn
  pivot_longer(`Nesting Male`:`Sneaker Male`,
               names_to = 'Pathway',
               values_to = 'Success Rate') |>
  #Graph it
  ggplot(aes(Pathway, `Success Rate`, fill = Pathway)) +
  #Makes the graph a bar graph
  geom_col() +
  labs(x = 'Pathway', 
       y = 'Times Mated',
       title = "The Number of times each Female spawned for each Visitation Pathway",) + #edits axis and title labels
  theme_classic() + #Sets the plot visuals
  #Adds a color scheme to the plot
  scale_fill_manual(values = c('Nesting Male' = '#065143', 'Satellite Male' = '#db5a42', 'Satellite Pass' = '#7A9B76','Sneaker Male' = '#1481BA')) +
  theme(plot.title = element_text(hjust = 0.5)) + #Moves title to the middle
  theme(legend.position = "none") # Removes all legends
  
pathways.success.rate.total
```

Now let's run some models

We'll run the Nesting Male and Satellite pass to Nesting Male pathways first since it seemed the best fit before.

```{r}
path.visits.norm <- behavior_sheet_cln |>
  #Normalize the different pathways by the total number of females that visited
  mutate(
    'nm.path' = (female_visits_by_nm)/(total_female_visits),
    'sat.path' = (female_visit_by_sat)/(total_female_visits),
    'sat.pass.path' = (female_visits_by_sat_passed_to_nm)/(total_female_visits),
    'snk.path' = (female_visit_by_snk)/(total_female_visits)
  )

path.visits.norm.rmv <- path.visits.norm |>
  #Removing rows where there were data mistakes
  filter(!(total_female_spawning > total_female_visits))


path.fem.spawning <- path.visits.norm.rmv |>
  brm(data = path.visits.norm.rmv,
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(total_female_spawning | trials(total_female_visits) ~ 1 + nm.path + sat.pass.path),
    prior = c(prior(normal(0, 2), class = Intercept)),
    iter = 4000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/path.fem.spawning') #saves it as an output

#Gives a plot of the model's predictions
pp_check(path.fem.spawning)
#Gives us the summary of the model
print(path.fem.spawning, digits = 4)
#Plots the Markov chains
plot(path.fem.spawning)
```

Let's create a model that will inform us about each pathway.

First we create the pathways

```{r}
pathway.spawns.modl <- behavior_sheet_cln |>
  #Remove nests where there are data inconsistencies
 filter_out((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm) > (total_spawns_nm_brought + total_spawns_nm_brought_sat_only_spawn)) |>
  filter_out((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn) > (total_spawns_sat_brought + total_spawns_sat_brought_sat_only_spawn)) |>
  filter_out((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn) > (total_spawns_sat_passed_to_nm + total_spawns_sat_passed_to_nm_but_sat_only_spawn))|>
  filter_out((female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn) > (total_spawns_snk_brought + total_spawns_snk_brought_snk_only_spawn))|>
   #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = (total_spawns_nm_brought + total_spawns_nm_brought_sat_only_spawn),
    'Satellite Male' = (total_spawns_sat_brought + total_spawns_sat_brought_sat_only_spawn),
    'Satellite Pass' = (total_spawns_sat_passed_to_nm + total_spawns_sat_passed_to_nm_but_sat_only_spawn),
    'Sneaker Male' = (total_spawns_snk_brought + total_spawns_snk_brought_snk_only_spawn)) |>
  #Move the pathways columns we just created into their own column for inputting into the model
  pivot_longer(cols = c(`Nesting Male`:`Sneaker Male`),
               names_to = 'Pathway',
               values_to = 'number.spawned')
```

Then we create the model that uses total spawns at a nest as the number of trials

```{r}
pathway.spawns.mdl.1 <- pathway.spawns.modl |>
  brm(
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(number.spawned | trials(total_spawns) ~ 0 + Pathway + (1|nest_name)),
    prior = c(prior(normal(0, 2), class = b)),
    iter = 4000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
      file = 'Data Analysis Outputs/pathway.spawns.mdl.1') #saves it as an output

#Gives a plot of the model's predictions
pp_check(pathway.spawns.mdl.1)
#Gives us the summary of the model
print(pathway.spawns.mdl.1, digits = 4)
#Plots the Markov chains
plot(pathway.spawns.mdl.1)
```

Now with visits

```{r}
pathway.visits.modl <- behavior_sheet_cln |>
  #Remove nests where there are data inconsistencies
  filter_out((female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm)/(total_female_spawning) > 1 ) |>
  filter_out((female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn)/(total_female_spawning) > 1) |>
  filter_out((female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn)/(female_visits_by_sat_passed_to_nm) > 1) |>
  filter_out((female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)/(total_female_spawning) > 1)|>
  #Create new variables that account for the pathway's proportion
  mutate(
    'Nesting Male' = (female_spawning_with_nm_brought_by_nm + female_spawning_with_only_sat_brought_by_nm),
    'Satellite Male' = (female_spawning_sat_brought + females_spawing_sat_brought_sat_only_spawn),
    'Satellite Male Pass' = (female_spawning_sat_passed_to_nm + female_spawning_sat_passed_to_nm_but_only_sat_spawn),
    'Sneaker Male' = (female_spawning_snk_brought + female_spawning_snk_brought_skn_only_spawn)
  ) |>
  #Move the pathways columns we just created into their own column for inputting into the model
  pivot_longer(cols = c(`Nesting Male`:`Sneaker Male`),
               names_to = 'Pathway',
               values_to = 'number.spawned')
  
```

Model

```{r}
pathway.spawns.mdl.nest <- pathway.visits.modl |>
  brm(
    family = binomial(link = 'logit'), #Tell model what type of distribution
    #Specifying the model
    formula = bf(number.spawned | trials(total_female_visits) ~ 0 + Pathway + (1|nest_name)),
    prior = c(prior(normal(0, 2), class = b)),
    iter = 4000, warmup = 1000, chains = 4, cores = 4, #Markov chain parameters
      seed = 4, #Determines which random numbers
    sample_prior = 'yes', #Allows us to run hypothesis tests of the different Pathways
      file = 'Data Analysis Outputs/pathway.spawns.mdl.nest') #saves it as an output

#Gives a plot of the model's predictions
pp_check(pathway.spawns.mdl.nest)
#Gives us the summary of the model
print(pathway.spawns.mdl.nest, digits = 4)
#Plots the Markov chains
plot(pathway.spawns.mdl.nest)

#This is how we are going to check if the Satellite pathways are statistically significantly different
sat.vs.satpath <- hypothesis(pathway.spawns.mdl.nest, 'PathwaySatellite Male = PathwaySatellite Male Pass', alpha = 0.05)
plot(sat.vs.satpath)
```

This hypothesis test tells us that the two satellite male pathways are not statistically significantly different from one another.

Now lets back-transform our data

```{r}
back.transformed <- data.frame('Estimate' = c(plogis(-0.7551),plogis(-3.8650), plogis(-3.6807), plogis(-6.2933)), row.names = c("Nesting Male", 'Satellite Male', "Satellite Male Pass", 'Sneaker Male'))

back.transformed.summary <- back.transformed |>
  mutate('l-95% CI' = c(plogis(-0.9013), plogis(-4.1906), plogis(-3.9848), plogis(-7.3479))) |>
  mutate('u-95% CI' = c(plogis(-0.6128), plogis(-3.5607), plogis(-3.3964), plogis(-5.4461)))
back.transformed.summary

```

So based on our models it seems like there is an effect on spawning for each of the pathways (albeit practically non-existent in some). Given our data it seems that for:

The probability of a female spawning when brought by the nesting male is around 31.9%

The probability of a female spawning when brought by the satellite male is around 2.05%

The probability of a female spawning when brought by the satellite male then passed to the nesting male is around 2.46%

The probability of a female spawning when brought by the sneaker male is 0.1%.

Since none of our 95% confidence intervals overlap with zero, we can assume that these effects are different from zero (again based on our data).

Thinking about this, it makes sense. As we saw in the graphs, the majority of female visits are by the nesting male, thus it would have a higher probability than the other pathways. It is interesting that the satellite male pathways are so low, however, the pathway when the satellite male passes the female to the nesting male being higher than when it's just the satellite male is super interesting, and is going to be something we're going to continue to look at. Overall, these visualizations and models help give us a foundation for what we're interested in!

\
This next chunk is the graph of our model's predictions

```{r}
#Create the predictions
preds <- predict_response(pathway.spawns.mdl.nest, terms = c('Pathway'))
#Plots the prediction
plot(ggpredict(pathway.spawns.mdl.nest, # Tells the data
               c("Pathway"), #Which term is on the x-axis
               type = "fixed", #it is fixed
               condition = c(total_female_visits = 10) #How many "trials to show"
     )) + 
  theme(legend.position = "none") + #Removes any legend
  theme_classic() + # Sets the overall theme
  labs(y = 'Probability of Spawning',
       title = 'Predicted Spawning Success for each pathway') + #Adjust axis and title
  theme(plot.title = element_text(hjust = 0.5)) #Moves title to the middle
```

This tells us some interesting information. Given 10 random trials, we would expected around 3 females to spawn when brought by the Nesting Male, zero for Sneaker males, and almost zero for the other types, this prediction makes sense. Since these pathways are counting when the Nesting Male and Satellite Male are openly spawning, but not all the sneak spawns from sneaker and satellites.

# Final notes

You are now caught up with our where our research stands! While our findings thus far can seem "disappointing" or "lack-luster" that is actually good! Nature is so complex, let alone animal behavior, thus any findings provide useful information! For our research, these models indicate that we need to adjust how we collect our behavior data. So far, we've only been working on total spawns of each pathway and not looking at how/what influences individual females (because this is going to be a lot of work, that's where you come in!). Additionally, while it is small, and not statistically significant, the difference between the satellite male and the satellite male pass pathway is promising! Maybe, with some adjustment to how we are quantifying our data, we can test if the different pathways matter. Which would be very interesting because that means the females are paying attention to the social context around them!

# Session info

Here is the session info!

```{r}
sess <- sessionInfo()
sess
```
